---
title: "ETC5513 - Assignment 1"
author: "Shruti Pillai"
format: 
    html: 
        toc: true
        theme: solar
    pdf: 
        toc: true
        theme: solar
---
```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(lubridate)
```

```{r reading data, include=FALSE}
raw_dataset <- read_csv("data/32544804_Shruti Pillai.csv")
```

## Introduction
The data used in this report is customer-level data which can aid us in understanding where to focus marketing efforts by showing us what actually drives purchasing behaviour. To optimise purchasing outcomes relative to marketing expenditure, it is essential to quantify the relationship between **marketing_spend**, **purchase_amount**, and customer engagement **(visits)**. Key patterns will be summarised using visualisations and tables in order to support interpretation. Therefore, this report will be presented as a reproducible quarto document which aims to support the research question.

## Research Question
In essence, the data presents us with three possible relationships to be investigated which can be summarised as one research question: How do **marketing_spend**, and customer engagement **(visits)** relate to **purchase_amount**, and does this relationship vary between different customer **segments**? 

## Dataset Introduction
```{r dataset inspection}
head(raw_dataset, 10)
glimpse(raw_dataset)
str(raw_dataset)
names(raw_dataset)
nrow(raw_dataset)
ncol(raw_dataset)
unique(raw_dataset$segment)
unique(raw_dataset$notes)
table(raw_dataset$notes, useNA = "ifany")
```


In this section, briefly describe your data (i.e. what the data is measuring or recording) in five sentences maximum using markdown.

To do this, you will need to clean your dataset. Include a description of this cleaning process (not the code) in the report. The code to clean the dataset must be included in the Quarto file, but not in the report.

### Dataset Description
Create a subsection that will report details about your data. You must include the size of the dataset such as the number of observations, variables and variable types. You must include a sentence that includes inline R code describing the number of variables and observations in your dataset.

Create two other summaries of your data, one as a plot and one as a table. Examples include counts of certain variables in groups, or summary statistics of numeric variables. Make sure these summaries are providing a broader context to the report and are not just arbitrary commentaries.

## Results
Using visualisations of the data, discuss the answer to your research questions. You must:

Create 2 figures maximum of your data that will help you answer your research question. Each figure must have a caption using the options inside the R code chunk. Create the figures using the ggplot2 package.
The figures must have captions, and must be referred to in the text of the results section.
As well as the figures, you must also include a paragraph describing these results and how they link back to your research question. Remember, you can use inline code to extract certain numbers automatically.
