---
title: "ETC5513 - Assignment 1"
author: "Shruti Pillai"
format: 
    html: 
        toc: true
        theme: solar
    pdf: 
        toc: true
        theme: solar
---
```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(lubridate)
```

```{r reading data, include=FALSE}
raw_dataset <- read_csv("data/32544804_Shruti Pillai.csv")
```

## Introduction
The data used in this report is customer-level data which can aid us in understanding where to focus marketing efforts by showing us what actually drives purchasing behaviour. To optimise purchasing outcomes relative to marketing expenditure, it is essential to quantify the relationship between **marketing_spend**, **purchase_amount**, and customer engagement **(visits)**. Key patterns will be summarised using visualisations and tables in order to support interpretation. Therefore, this report will be presented as a reproducible quarto document which aims to support the research question.

## Research Question
In essence, the data presents us with three possible relationships to be investigated which can be summarised as one research question: How do **marketing_spend**, and customer engagement **(visits)** relate to **purchase_amount**, and does this relationship vary between different customer **segments**? 

## Dataset Introduction
```{r dataset inspection, include=FALSE}
head(raw_dataset, 10)
glimpse(raw_dataset)
str(raw_dataset)
names(raw_dataset)
nrow(raw_dataset)
ncol(raw_dataset)
unique(raw_dataset$segment)
unique(raw_dataset$notes)
table(raw_dataset$notes, useNA = "ifany")
anyDuplicated(raw_dataset)
```

```{r dataset cleaning, include=FALSE}
#Standardised notes section and checked if changes were correctly applied.

cleaned_dataset <- raw_dataset |>
    mutate(
        notes = na_if(notes, "N/A")
    )
table(cleaned_dataset$notes, useNA = "ifany")

#Standardised date format and checked if changes were correctly applied. Since this category won't be used in the analysis, it hasn't been perfectly cleaned.

cleaned_dataset <- cleaned_dataset |>
    mutate(
        signup_date = parse_date_time(signup_date, orders = c("ymd", "B d, Y", "dmy", "b d, Y"))  
    |> as.Date()
    )
table(cleaned_dataset$signup_date, useNA = "ifany")

#Converted visits from numeric to integer and checked if changes were correctly applied.

cleaned_dataset <- cleaned_dataset |>
    mutate(
        visits = as.integer(visits)
    )
str(cleaned_dataset$visits)

#Overall check
str(cleaned_dataset)
```

Investigation shows that the data consists of customer records that assign each individual their own unique **customer_id**. Each customer has their **marketing_spend** (marketing expenditure), **purchase_amount**, **segment**, **signup_date**, **notes** on interactions and number of **visits** (engagement count) recorded. The segment variable splits customers into three different groups which allows for comparison: **consumer**, **smb** and **enterprise**. The signup_date and notes variables, while not relevant to this report, indicate when i customer signed up and sore simple notes about interactions with the customer respectively. To summarise, the dataset contains on how marketing and engagemnt relate to purchasing behaviour amongst individuals. 

#### Cleaning Process Summary
In order to clean the dataset, the following process was implemented: 

1. The data was thorougly inspected using functions such as `head()`, `glimpse()`, `str()`, `names()`, `nrow()`, `ncol()`, `unique()`, and `table()` to understand the structure of the data, identify any duplicates or inconsistencies.

2. From this, **notes** and **signup_date** were found to be inconsistent and needed to be cleaned, while **visits**, not necessarily needing to be altered, would be smoother if changed from numeric to integer.

3. The following steps were saved into a new dataset called `cleaned_dataset` to ensure that the original data was preserved and could be referred back to if needed:

    a. **notes** were cleaned by standardising the entries to NA where they were previously recorded as a mix of "N/A" and the expected NA. This was done using the `na_if()` function.

    b. **signup_date** was cleaned by standardising the date format using the `parse_date_time()` function from the lubridate package. It allowed for multiple formats to be parsed and converted to a consistent date format. There were around 3 entries that couldn't be parsed but were left as is since the variable isn't a category of interest.

    c. **visits** were converted from numeric to integer using the `as.integer()`. This was done to ensure that the data type properly reflected the nature of the variable.



## Results
Using visualisations of the data, discuss the answer to your research questions. You must:

Create 2 figures maximum of your data that will help you answer your research question. Each figure must have a caption using the options inside the R code chunk. Create the figures using the ggplot2 package.
The figures must have captions, and must be referred to in the text of the results section.
As well as the figures, you must also include a paragraph describing these results and how they link back to your research question. Remember, you can use inline code to extract certain numbers automatically.
